### 2019-06-30 draft.tex
more dummy text

>>>>

The plasma physics branch of Physics is broad. From life-changing
industrial applications like the plasma processing of integrated
circuits - an essential step of the manufacturing process of the
tiny integrated circuits -, passing through the development of the
telecommunications via radio-frequency, to day-to-day devices
like fluorescent lamps \cite{bell,gurnett2017,chen}. Another
important application of plasmas on Earth

% The first fact that a plasma physics student learns, before even
% understanding the definition of plasma, is that most of the baryonic
% matter in the universe is in the plasma state. This \emph{most} is
% usually estimated to be around $99\%$ \cite{bitt,gurnett2017}. If this
% high estimate has something to do with the fact that our probes and
% telescopes can only directly detect ionized matter, because it emits
% radiation, \cite{chen} it is still a case to be studied.

% In order to understand what is a plasma, one must have to be aware of
% the differences between the dynamics of neutral and fully ionized gases.
% In a neutral gas, assuming that the influence of gravity can be neglected,
% the dynamics is dictated by collisions between two components that are
% very close to each other \cite{klimon}. A fully ionized gas, on the other
% hand, is only composed of charged particles, which are sources of electric
% fields. Besides that, the motion of charged particles generates electric
% currents, that will induce magnetic fields. In such arrangement, one charge
% can interact with several other charges at a considerable distance trough
% the action of the self-generated electric and magnetic fields. Therefore, a
% partially ionized gas is considered a plasma when its degree of ionization
% is such that its dynamics is dominated by the electric and magnetic forces
% of the charged particles. Under such definition, a fully ionized gas is, of
% course, also considered a plasma.

% Plasmas can be classified according with their overall neutrality as
% \emph{neutral} when the density of electrons and ions are the same, and
% \emph{non-neutral} when the densities of the negatively and positively
% charged particles are different, which includes the case of one component
% plasma.

% In the beginning of this chapter it was mentioned a fundamental
% characteristic of a plasma: the \emph{collective processes}.
% Indeed, an ionized gas cannot be called a plasma if its dynamics
% is not predominantly collective.

% In order to define collective processes, one should have in mind that
% a plasma is a gas that has most of its components in an ionized state.
% For instance, let us consider a hydrogen gas that was heated to the
% point where the electrons were separated from their respective ions.
% What we have now is gas that is composed of charged particles that
% are source of electric fields. Also, the motion of the electrons
% generates electric currents that induce magnetic fields.

On Earth, the plasma is considered an exotic state of matter, associated to the
glowing gas inside fluorescent and neon lamps, microchip manufacturing, research
laboratories, and the promise of cleaner and technically infinite energy source
when scientists and engineers manage to succeed at obtaining a controlled, and
energetically efficient, thermonuclear fusion. Naturally occurring plasmas in our
everyday life are rare, restricted to the occurrence of lightning during storms
and beautiful auroras that are spotted at high latitudes. In the interplanetary
and interstellar space, however, the situation is the opposite. Starting in the
outer layers of Earth's atmosphere, passing through the solar wind in the
interplanetary space, to the stellar interiors and atmospheres, accretion disks,
and molecular clouds it is estimated that something around $99 \sim 99.9\%$ of
the observable matter is in the plasma state \cite{may}.

What defines the plasma state in an ionized gas, is the presence of collective
behavior, which is a consequence of the long-range action of the electric and
magnetic forces that are generated by the charged particles and their movement.
There are other important properties, like Debye shielding and quasineutrality,
that are well described in introductory books \cite{chen,bitt,gurnett2017} and
will not be addressed here. Beyond these fundamental properties, plasmas can
be found in an immense variety of configurations, and this is notably true for
the plasma in space.
